### 2024-04-02 scripts/01-download_data.qmd
update

#### Preamble ####
# Purpose: Downloads and saves the data from from the Womens´s Pockets are Inferior paper by Jan Diehm & Amber Thomas. 
# Author: Hari Lee Robledo
# Date: April 2 2023 
# Contact: hari.leerobledo@mail.utoronto.ca 
# License: MIT
# Pre-requisites: None

```{r}
#### Workspace setup ####
#install.packages()
library(tidyverse)
library(readr)
```

``` {r}
#### Download data ####

# Corrected URL to the raw CSV file
url <- "https://raw.githubusercontent.com/the-pudding/data/725dc18a59a055891dee25e85652f416b0065bdd/pockets/measurements.csv"

# Read the CSV file
raw_pockets_data <- read_csv(url)

```
```{r}
#### Save data ####
write_csv(
  x = raw_pockets_data,
  file = "data/raw_data/raw_pockets_data.csv"
)
```